Automated topic naming to support cross-project analysis of software maintenance activities

### 2011-03-24 msr/msr2011.tex
justification for validity of eval data     - [X] the authors need to provide more justification for the      validity of the evaluation data, preferably include annotators      outside the authors, and evaluate agreement. What justification      do the authors have that the data is accurate?           - response: the annotators have much experience in software development             and did inspect some of the labels that the other             annotators were producing. Unfortunately we lack the data             to run inter-rater reliability but this was already             pointed out in t

\begin{abstract}
% %\todo[inline]{Topics and LDA sooner, Topics/concept location is a known
Researchers have employed a variety of techniques to extract underlying
topics that relate to software development artifacts.  Typically, these
techniques use unsupervised machine-learning algorithms to suggest
candidate word-lists.  However, word-lists are
difficult to 
interpret in the absence of meaningful summary labels.  Current topic modeling 
techniques assume manual labelling and do not use domain-specific knowledge
to improve, contextualize, or describe results for the developers.  
We propose a solution: \emph{automated labelled topic extraction}.
  Topics are extracted 
using  Latent Dirichlet Allocation (LDA) from commit-log comments recovered
from source control systems such as CVS and BitKeeper.  
These topics are given labels from a generalizable
cross-project taxonomy, consisting of non-functional
requirements. Our approach was evaluated with experiments and case studies
on two large-scale RDBMS projects: MySQL and MaxDB.  The case studies
show that labelled topic extraction can produce appropriate,
context-sensitive labels relevant to these projects, which provides fresh
insight into their evolving software development activities.
\end{abstract}

@@

%We first introduces labelled topic extraction, and then uses that technique to analyze development activity in two large-scale examples.
We begin by discussing related work in Section \ref{sec:related}.
Next, we describe how we generated our data (Section \ref{sec:wordlist}). For unsupervised classification (Section \ref{sec:unsuplabelling}), we begin by creating word-lists to signify when a topic matches an NFR label. We then apply our classifier and analyze the results. %and how we created the word-lists used in unsupervised classification. 
%We show that labels with their topics can be learned and used to classify other data-sets, either without training 
%Our goal is to label a given topic with either an NFR from our taxonomy,
 %Section \ref{sec:unsuplabelling} presents our technique for using word-lists to do unsupervised classification, and describes the results.
In Section \ref{sec:suplabelling}, we manually annotate the topics, and use those annotations as training data for supervised classification.  
%We then present visualizations of named topics and their trends over time to aid communication and analysis. 
To demonstrate an application of labelled topic extraction, we use an exploratory case study of two open source database systems to show how named topics can be compared between projects  (Section \ref{sec:analysis}). 
The paper concludes with a discussion of limitations (Section \ref{sec:limit}), and future work.

@@

\section{Study design and execution}
Figure \ref{fig:process} gives an outline of methodology.
We began by gathering source data and creating topic models. For unsupervised labelling, we generated three sets of word-lists as signifiers for NFRs.
%e created labels using a well-known NFR taxonomy. 
In supervised learning, we trained our data with manual annotations in order to match topics with NFRs. Finally, these topics were used to analyze the role of NFRs in software maintenance.

@@

%\noindent \textbf{Creating a validation corpus} --- 
\subsubsection{Creating a validation corpus}
To evaluate both unsupervised and supervised classification, we
created a validation set of manually labelled topics. For MySQL 3.23
and MaxDB 7.500, we (the first two authors) annotated each extracted topic in each period with
the six NFR labels listed above.
Annotators did not annotate each other's annotations but some brief
inspection of annotations was used to confirm that the annotators were
acting similarly.
We looked at each period's topics, and assessed what the data --- consisting of the frequency-weighted word lists and messages --- suggested was the topic for that period. 
We were able to pinpoint the appropriate label using auxiliary information as well, such as the actual revisions and files that were related to the topic being annotated.
For example, for the MaxDB topic consisting of a message ``exit() only
used in non NPTL LINUX Versions'', we tagged that topic
\emph{portability}. 
Given the top-level annotations of \emph{none}, \emph{portability},
\emph{efficiency}, \emph{reliability}, \emph{functionality},
\emph{usability}, and \emph{maintainability}, the annotators annotated each topic
with the relevant label and sometimes we used finer-grained
annotations that would be aggregated up to one of these higher-level labels.
%\todo[inline]{this confused people. Is it truly unsupervised? Yes, but it needs to be clearer.}
%We compared against this data-set, but we also used the data-set for our supervised machine learning based topic classification. 
% A topic may have more than one matching keyword. % ok what?
%\todo[inline]{inter-rater reliability }
% \todo[inline]{There should be a section here which is about manual annotation. It is really important and it is required by both analyses. Plus we can tell them how.}

@@

\subsection{Unsupervised labelling}
\label{sec:unsuplabelling}

@@

Using our three word-lists (\textsf{exp1}, \textsf{exp2}, \textsf{exp3}), we labeled our topics with an NFR where there was a match between a word in the list and the same word somewhere in the distribution of words that constitute the topic.
A \emph{named topic} is a topic with a matching NFR label. Unnamed
topics occur where there is no such match, which may indicate either a lack of precision, or simply that this topic is not associated with non-functional requirements.
All experiments were run on MaxDB 7.500 and MySQL 3.23 data. LDA
extracted 20 topics per period for each project.
% Is this good enough
This labelling is unsupervised because the corpus is not derived from 
the project being labelled, and we did not label the project's topics
ourselves for a training set. The motivation behind this technique is that
because most software often addresses similar issues, we can use the
domain knowledge of software to label relevant topics.
%\todo[inline]{Neil does this explain the unsupervised aspect well enough?}

@@

\subsubsection{Analysis of the unsupervised labelling} %--- %Based
                                %on the labels, and our manual topic
                                %labelling, 
%\noindent \textbf{Analysis of the unsupervised labelling} --- %Based on the labels, and our manual topic labelling, 
%we compared the results of the unsupervised word matching approach. 
For each quality we assessed whether unsupervised labels matched the manual annotations. %To be clear, the manual annotations were not used to train the labeling process.
As described in \ref{sec:wordlist} we used both ROC and F-1 measures to evaluate the performance of the classification.
Figure \ref{fig:maxdb-unsup-results} shows our ROC results for MaxDB and MySQL. We describe F-1 results in the text below.

@@

\begin{figure*}
  \centering
 \includegraphics[width=0.8\textwidth]{figures/unsupervised-bar}
 \caption{Performance, ROC values (range: 0-1), of unsupervised topic labelling for
   each NFR and per word-list. Dashed line indicates the performance of a random classifier. This graph shows how well the
   unsupervised topic labelling matched our manual annotations.}

@@

Very rarely did \textsf{exp2} and \textsf{exp3} (unsupervised word matching) ever perform as well as the supervised machine learners. 
For MaxDB, \textit{reliability} was slightly better detected using the static word list of \textsf{exp2}. 
In general, the machine learners and \textsf{exp3} did better than \textsf{exp2} for both MaxDB and MySQL. 
For both MySQL and MaxDB \textit{usability} was better served by \textsf{exp2}. 
\textit{Usability} was a very infrequent label, however, which made it difficult to detect in any case.

The unsupervised labelling had difficulty distinguishing between common labels and infrequent labels. 
The learners would occasionally mislabel a topic deserving of an infrequent label with a more common label.
The word-lists for \emph{correctness} tended to be too lengthy, non-specific and broad, especially if WordNet words were used, since the NFRs are typically loosely defined concepts in common parlance.

@@

If the techniques used in \textsf{exp2} and \textsf{exp3} were combined with the supervised techniques, we could reduce the training effort by boosting training sets with topics classified with the unsupervised techniques.
Both Naive Bayesian learners and the word-list approaches were computationally efficient.  
%These results are promising, because the accuracy of the techniques is good enough to be useful, while the run-time performance is still cheap enough to execute to be feasible as an automated or semi-automated method of labelling topics by their software qualities.
These results are promising because they indicate that these
techniques are accurate enough to be useful while still maintaining
acceptable run-time performance.

@@

\emph{Internal validity} -- %includes inter-rater reliability and our lack of rival explanation building. 
We improved internal validity by trying to correlate and explain the behaviours observed in the analysis with the historical records of the projects.
We did not attempt to match our results to any particular model or theory.

@@

\section{Conclusions}
This paper presented a cross-project data mining technique,
\emph{labelled topic extraction}. %The technique leveraged a
%non-functional requirements (NFRs) taxonomy to mine and label latent
%software development topics. 
Previous topic analysis research produced project-specific
topics that needed to be manually labelled.
%, but topics without labels
%are difficult to interpret, and manually labelling them is
%costly. Furthermore, previous work was project-specific and thus incomparable. 
To improve on this, we leveraged software engineering standards, specifically the ISO quality taxonomy, to produce a method of partially-automated (supervised) and fully-automated (unsupervised) topic labelling.
Since the word-list technique is not project-specific, we used it to compare two distinct projects, where we showed our technique produced interesting insight into maintenance activity. %topics and artifacts from project histories.
% Our contributions include:
% \begin{itemize}
% \item We provided an unsupervised method of topic labelling (word-lists);
% %\item Our unsupervised method can be applied to projects without training;
% \item We demonstrated a supervised method of labelling topics by a single NFRs (machine learning);
% \item We demonstrated a supervised method of labelling topics by multiple NFRs (multi-label machine learning);
% \item We provided a method of cross-project analysis via topic labelling;
% \item We demonstrated a method of visualizing these topics across time, which we used to analyze maintenance activities.
% \end{itemize}

We validated our topic labelling techniques using multiple experiments. % on unsupervised learners, supervised and multi-label learners. 
We first conducted unsupervised labelling using word-lists. 
Our next approach was supervised, using single-label and multi-label learners. 
Both kinds of learners performed well with average
%area under the 
ROC values between $0.6$ and $0.8$. These results, along with the efficient performance of our learners, demonstrate that labelled topic extraction is a promising approach for understanding the occurrence of non-functional requirements in software projects. 
%Although it may seem like common knowledge that maintenance is a concern as a project matures, our technique was able to show this independently.
%We showed that NFRs are often trending topics in commit comments
%, and that NFRs are quite common as developer topics.
%; and that there are efficient methods of automating topic labelling.

>>>>

\begin{abstract}
% %\todo[inline]{Topics and LDA sooner, Topics/concept location is a known
Researchers have employed a variety of techniques to extract underlying
topics that relate to software development artifacts.  Typically, these
techniques use semi-unsupervised machine-learning algorithms to suggest
candidate word-lists.  However, word-lists are
difficult to 
interpret in the absence of meaningful summary labels.  Current topic modeling 
techniques assume manual labelling and do not use domain-specific knowledge
to improve, contextualize, or describe results for the developers.  
We propose a solution: \emph{automated labelled topic extraction}.
  Topics are extracted 
using  Latent Dirichlet Allocation (LDA) from commit-log comments recovered
from source control systems such as CVS and BitKeeper.  
These topics are given labels from a generalizable
cross-project taxonomy, consisting of non-functional
requirements. Our approach was evaluated with experiments and case studies
on two large-scale RDBMS projects: MySQL and MaxDB.  The case studies
show that labelled topic extraction can produce appropriate,
context-sensitive labels relevant to these projects, which provides fresh
insight into their evolving software development activities.
\end{abstract}

@@

%We first introduces labelled topic extraction, and then uses that technique to analyze development activity in two large-scale examples.
We begin by discussing related work in Section \ref{sec:related}.
Next, we describe how we generated our data (Section \ref{sec:wordlist}). For semi-unsupervised classification (Section \ref{sec:unsuplabelling}), we begin by creating word-lists to signify when a topic matches an NFR label. We then apply our classifier and analyze the results. %and how we created the word-lists used in unsupervised classification. 
%We show that labels with their topics can be learned and used to classify other data-sets, either without training 
%Our goal is to label a given topic with either an NFR from our taxonomy,
 %Section \ref{sec:unsuplabelling} presents our technique for using word-lists to do unsupervised classification, and describes the results.
In Section \ref{sec:suplabelling}, we manually annotate the topics, and use those annotations as training data for supervised classification.  
%We then present visualizations of named topics and their trends over time to aid communication and analysis. 
To demonstrate an application of labelled topic extraction, we use an exploratory case study of two open source database systems to show how named topics can be compared between projects  (Section \ref{sec:analysis}). 
The paper concludes with a discussion of limitations (Section \ref{sec:limit}), and future work.

@@

\section{Study design and execution}
Figure \ref{fig:process} gives an outline of methodology.
We began by gathering source data and creating topic models. For semi-unsupervised labelling, we generated three sets of word-lists as signifiers for NFRs.
%e created labels using a well-known NFR taxonomy. 
In supervised learning, we trained our data with manual annotations in order to match topics with NFRs. Finally, these topics were used to analyze the role of NFRs in software maintenance.

@@

%\noindent \textbf{Creating a validation corpus} --- 
\subsubsection{Creating a validation corpus}
To evaluate both semi-unsupervised and supervised classification, we
created a validation set of manually labelled topics. For MySQL 3.23
and MaxDB 7.500, we (the first two authors) annotated each extracted topic in each period with
the six NFR labels listed above.
Annotators did not annotate each other's annotations but some brief
inspection of annotations was used to confirm that the annotators were
acting similarly.
We looked at each period's topics, and assessed what the data --- consisting of the frequency-weighted word lists and messages --- suggested was the topic for that period. 
We were able to pinpoint the appropriate label using auxiliary information as well, such as the actual revisions and files that were related to the topic being annotated.
For example, for the MaxDB topic consisting of a message ``exit() only
used in non NPTL LINUX Versions'', we tagged that topic
\emph{portability}. 
Given the top-level annotations of \emph{none}, \emph{portability},
\emph{efficiency}, \emph{reliability}, \emph{functionality},
\emph{usability}, and \emph{maintainability}, the annotators annotated each topic
with the relevant label and sometimes we used finer-grained
annotations that would be aggregated up to one of these higher-level labels.
%\todo[inline]{this confused people. Is it truly unsupervised? Yes, but it needs to be clearer.}
%We compared against this data-set, but we also used the data-set for our supervised machine learning based topic classification. 
% A topic may have more than one matching keyword. % ok what?
%\todo[inline]{inter-rater reliability }
% \todo[inline]{There should be a section here which is about manual annotation. It is really important and it is required by both analyses. Plus we can tell them how.}

@@

\subsection{Semi-unsupervised labelling}
\label{sec:unsuplabelling}

@@

Using our three word-lists (\textsf{exp1}, \textsf{exp2}, \textsf{exp3}), we labeled our topics with an NFR where there was a match between a word in the list and the same word somewhere in the distribution of words that constitute the topic.
A \emph{named topic} is a topic with a matching NFR label. Unnamed
topics occur where there is no such match, which may indicate either a lack of precision, or simply that this topic is not associated with non-functional requirements.
All experiments were run on MaxDB 7.500 and MySQL 3.23 data. LDA
extracted 20 topics per period for each project.
% Is this good enough
This labelling is \emph{semi-unsupervised} because the corpus is not derived from 
the project being labelled, and we did not label the project's topics
ourselves for a training set. The motivation behind this technique is that
because most software often addresses similar issues, we can use the
domain knowledge of software to label relevant topics.
%\todo[inline]{Neil does this explain the unsupervised aspect well enough?}

@@

\subsubsection{Analysis of the semi-unsupervised labelling} %--- %Based
                                %on the labels, and our manual topic
                                %labelling, 
%\noindent \textbf{Analysis of the unsupervised labelling} --- %Based on the labels, and our manual topic labelling, 
%we compared the results of the unsupervised word matching approach. 
For each quality we assessed whether semi-unsupervised labels matched the manual annotations. %To be clear, the manual annotations were not used to train the labeling process.
As described in \ref{sec:wordlist} we used both ROC and F-1 measures to evaluate the performance of the classification.
Figure \ref{fig:maxdb-unsup-results} shows our ROC results for MaxDB and MySQL. We describe F-1 results in the text below.

@@

\begin{figure*}
  \centering
 \includegraphics[width=0.8\textwidth]{figures/unsupervised-bar}
 \caption{Performance, ROC values (range: 0-1), of semi-unsupervised topic labelling for
   each NFR and per word-list. Dashed line indicates the performance of a random classifier. This graph shows how well the
   semi-unsupervised topic labelling matched our manual annotations.}

@@

Very rarely did \textsf{exp2} and \textsf{exp3} (semi-unsupervised word matching) ever perform as well as the supervised machine learners. 
For MaxDB, \textit{reliability} was slightly better detected using the static word list of \textsf{exp2}. 
In general, the machine learners and \textsf{exp3} did better than \textsf{exp2} for both MaxDB and MySQL. 
For both MySQL and MaxDB \textit{usability} was better served by \textsf{exp2}. 
\textit{Usability} was a very infrequent label, however, which made it difficult to detect in any case.

The semi-unsupervised labelling had difficulty distinguishing between common labels and infrequent labels. 
The learners would occasionally mislabel a topic deserving of an infrequent label with a more common label.
The word-lists for \emph{correctness} tended to be too lengthy, non-specific and broad, especially if WordNet words were used, since the NFRs are typically loosely defined concepts in common parlance.

@@

If the techniques used in \textsf{exp2} and \textsf{exp3} were combined with the supervised techniques, we could reduce the training effort by boosting training sets with topics classified with the semi-unsupervised techniques.
Both Naive Bayesian learners and the word-list approaches were computationally efficient.  
%These results are promising, because the accuracy of the techniques is good enough to be useful, while the run-time performance is still cheap enough to execute to be feasible as an automated or semi-automated method of labelling topics by their software qualities.
These results are promising because they indicate that these
techniques are accurate enough to be useful while still maintaining
acceptable run-time performance.

@@

\emph{Internal validity} -- %includes inter-rater reliability and our lack of rival explanation building. 
We improved internal validity by trying to correlate and explain the behaviours observed in the analysis with the historical records of the projects.
We did not attempt to match our results to any particular model or
theory.
Also we did not have enough manpower to assess inter-rater reliability.

@@

\section{Conclusions}
This paper presented a cross-project data mining technique,
\emph{labelled topic extraction}. %The technique leveraged a
%non-functional requirements (NFRs) taxonomy to mine and label latent
%software development topics. 
Previous topic analysis research produced project-specific
topics that needed to be manually labelled.
%, but topics without labels
%are difficult to interpret, and manually labelling them is
%costly. Furthermore, previous work was project-specific and thus incomparable. 
To improve on this, we leveraged software engineering standards, specifically the ISO quality taxonomy, to produce a method of partially-automated (supervised) and fully-automated (semi-unsupervised) topic labelling.
Since the word-list technique is not project-specific, we used it to compare two distinct projects, where we showed our technique produced interesting insight into maintenance activity. %topics and artifacts from project histories.
% Our contributions include:
% \begin{itemize}
% \item We provided an unsupervised method of topic labelling (word-lists);
% %\item Our unsupervised method can be applied to projects without training;
% \item We demonstrated a supervised method of labelling topics by a single NFRs (machine learning);
% \item We demonstrated a supervised method of labelling topics by multiple NFRs (multi-label machine learning);
% \item We provided a method of cross-project analysis via topic labelling;
% \item We demonstrated a method of visualizing these topics across time, which we used to analyze maintenance activities.
% \end{itemize}

We validated our topic labelling techniques using multiple experiments. % on unsupervised learners, supervised and multi-label learners. 
We first conducted semi-unsupervised labelling using word-lists. 
Our next approach was supervised, using single-label and multi-label learners. 
Both kinds of learners performed well with average
%area under the 
ROC values between $0.6$ and $0.8$. These results, along with the efficient performance of our learners, demonstrate that labelled topic extraction is a promising approach for understanding the occurrence of non-functional requirements in software projects. 
%Although it may seem like common knowledge that maintenance is a concern as a project matures, our technique was able to show this independently.
%We showed that NFRs are often trending topics in commit comments
%, and that NFRs are quite common as developer topics.
%; and that there are efficient methods of automating topic labelling.
